\documentclass{rudin}
\usepackage[]{hyperref}

\title{Computing $F$}
\author{Robert Dougherty-Bliss}
\date{\today}

\begin{document}

\maketitle

We want to evaluate the function
\begin{equation*}
    F(a, b) = 2 \sum_{k \geq 0} \frac{1}{(k + 2)! (3 + b / a)^{\overline{k}} a^k}
\end{equation*}
in the nice case where $b / a = m - 1/2$ for a nonnegative integer $n$. To this
end, let
\begin{equation*}
    S(m, z) = \sum_{k \geq 0} \frac{(k + m + 2)!}{(k + 2)! (2k + 2m + 4)!} z^k.
\end{equation*}
From the arguments on page 11 of \url{https://arxiv.org/pdf/2004.00090.pdf},
\begin{equation*}
    F(a, b) = \frac{2(4 + 2m)!}{(m + 2)!} S(m, 4 / a).
\end{equation*}
so we just need to evaluate $S(m, z)$.

Shift the summation index forward by $2$:
\begin{equation*}
    S(m, z) = z^2 \sum_{k \geq 2} \frac{(k + m)!}{k! (2k + 2m + 2)!}
\end{equation*}
Note that $(k + m)! / k! = (k + m)^{\underline{m}} = (k + m)(k + m - 1) \cdots
(k + 1)$, so
\begin{equation*}
    S(m, z) = z^{-2} \sum_{k \geq 2} \frac{(k + m)^{\underline{m}}}{(2k + 2m + 2)!} z^k.
\end{equation*}
If we now shift the summation index forward by $m$, we get
\begin{equation*}
    S(m, z) = z^{-2} z^{-m} \sum_{k \geq m + 2} \frac{k^{\underline{m}}}{(2k)!} z^k.
\end{equation*}
Now we're in luck, because we ``know''
\begin{equation*}
    \sum_{k \geq m + 2} \frac{1}{(2k)!} z^k;
\end{equation*}
it is $\cosh \sqrt{z}$ minus finitely many terms.

It is easy to check that, for any polynomial $P(k)$,
\begin{equation*}
    \sum_{k \geq 0} P(k) a_k z^k = P(zD) \sum_{k \geq 0} a_k z^k,
\end{equation*}
where $D$ is the differentiation operator. Thus,
\begin{align*}
    S(m, z) &= z^{-m - 2} (zD)^{\underline{m}} \sum_{k \geq m + 2} \frac{z^k}{(2k)!} \\
            &= z^{-m - 2} (zD)^{\underline{m}} (\cosh \sqrt{z} - \sum_{0 \leq k \leq m + 1} \frac{z^k}{(2k)!}).
\end{align*}

This is about the best that we should hope for as a closed-form. But for fixed
$m$, we can ask a computer to do some evaluations for us. For example,
\begin{equation*}
    S(2, z) =
{\frac {1}{{z}^{2}} \left( -{\frac {1}{4}\sinh \left( \sqrt {z} \right) {z}^{-{\frac{3}{2}}}}+{\frac {1}{4\,z}\cosh \left( \sqrt {z}
 \right) }-{\frac{1}{12}}-{\frac {z}{120}} \right) }
\end{equation*}
If we take $a = 4$ with $m = 2$, then $b = 6$, and we get
\begin{align*}
    F(4, 6) &= \frac{2 \cdot 8!}{4!} S(2, 4 / 4) \\
            &= 3360 ((\cosh 1 - \sinh 1) / 4 - 11 / 120) \\
            &= \frac{840}{e} - 308.
\end{align*}

Maple programs to do this exist here: \url{https://rwdb.xyz/files/programs/computeF}
\end{document}
